---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

# title: "Helping You Write Academic Papers in R using Texevier"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: TRUE
Entry1: "COVID-19 Vaccine Lottery Field Experiment"
Entry2: "\\textbf{Cassandra Pengelly | 20346212}" # textbf for bold
Entry3: "Professor R. Burger"
Uni_Logo: Tex/logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
Logo_width: 0.9 # If using a logo - use this to set width (size) of image
Entry4: "Behavioural Economics 871 Essay"
Entry5: "15 October 2021 | Word Count: 2970"
# Entry6: ""
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
# AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
# Author1: "Nico Katzke^[__Contributions:__  \\newline _The authors would like to thank no institution for money donated to this project. Thank you sincerely._]"  # First Author - note the thanks message displayed as an italic footnote of first page.
# Ref1: "Prescient Securities, Cape Town, South Africa" # First Author's Affiliation
# Email1: "nfkatzke\\@gmail.com" # First Author's Email address
# 
# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE
# 
# # Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: TRUE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
# abstract: |
  # Abstract to be written here. The abstract should not be too long and should provide the reader with a good understanding what you are writing about. Academic papers are not like novels where you keep the reader in suspense. To be effective in getting others to read your paper, be as open and concise about your findings here as possible. Ideally, upon reading your abstract, the reader should feel he / she must read your paper in entirety.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
```
<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->
\newpage

# Introduction \label{Introduction}

Coronavirus disease 2019 (COVID-19) has caused the largest public health crisis, and economic disaster of the 21st century so far [@bad p.471]. According to the @who, COVID-19 has directly resulted in 4,859,277 deaths worldwide^[As of 13 October 2021], and its impact on the global economy has been severe [@bank]. Policymakers are in urgent need of evidence-based strategies to contain the pandemic. As @immun[p.583] report, one critical mechanism through which epidemics are controlled is herd immunity. Herd immunity arises when a sufficiently large proportion of the population achieves individual immunity to an infectious disease such that the transmission chain of the disease is halted [@bad p.471]. One method of establishing herd immunity is through vaccination programs [@immun p.583]. Empirical evidence shows that vaccination strategies could be critical tools in subduing COVID-19. For example, @erad find COVID-19 vaccines to be effective in Jakarta, Indonesia. Leaders of other countries have also implemented vaccine programs and the @who reports that a total of 6,364,021,792 vaccine doses have been administered globally^[As of 10 October 2021].

In line with the international approach, South Africa has issued vaccines in addition to social distancing and lockdown measures. The South African government has stated that it aims to have 67% of the population vaccinated by the end of 2021 [@herd]. However, many South Africans are hesitant to get the COVID-19 vaccine; the @stat records that only 25% of South Africa's adult population has been fully vaccinated as of 11 October 2021. @cram compiled a report, based on data from wave 5 of the NIDS-CRAM survey, and find that that around 20% of South Africans are concerned that COVID-19 vaccines are not safe. The report concluded that a significant portion of South Africans still have to be convinced to take the vaccine^[A sentiment report from @report shows an interesting break down of different communities' beliefs surrounding the vaccine] [@cram]. While there are different ways to encourage vaccine uptake, such as mandatory vaccination or lump-sum transfers, insights from behaviourial economics could provide a more cost-effective solution: a vaccine lottery. 

This essay^[This essay was written in R using the package Texevier by @Texevier] proposes a field experiment to investigate whether different vaccine lotteries -- standard, regret and referral -- could improve vaccination rates in South Africa. This essay is structured as follows. Section \ref{lit} reviews the relevant literature on behavioural economics and health incentives. Section \ref{context} elaborates on the South African context. Section \ref{design} describes the design of the experiment and outlines the three types of treatment groups. Section \ref{treatment} discusses how the treatment will be administered and the data collection process. Section \ref{pre} gives a pre-analysis plan, and the final section (\ref{con}) concludes. 

# Behavioural Economics and Health Incentives  \label{lit} 

Health professionals and policymakers are increasingly turning to behavioural economics to understand how people make health decisions and how behavioural insights can be used to improve public health outcomes [@health p.1]. While neoclassical economics assumes that people are perfectly rational agents, behavioural economics relaxes this assumption and uses psychology and economic theory to create more realistic models of human decision-making [@rabin]. As Khaneman and Tversky show, people are subject to certain biases and often make use of heuristics in their decision-making process, which can lead to predictable errors in judgment [@khan \& @fast \& @prospect]. Behavioural economics literature investigates how these biases can be combated to improve welfare outcomes. @nudge introduced the idea of a nudge^[Nudge: an intervention that alters behaviour towards a desired action. In order for an intervention to qualify as a nudge, it should be cheap and easy to avoid [@nudge].] as a way to guide people to make better choices. For example, changing the default option for organ donation to be opt-in as opposed to explicit consent could benefit potential donors (who were deterred by the registration process) and save more lives [@nudge p.176].

A similar type of nudge can be applied to flu vaccines. A study conducted by @opt found that vaccination rates increased by 36% under an opt-in default than under an opt-out condition. @flu [p.9] proposes other interventions for promoting vaccinations, e.g. encouraging people to plan the time and location they will receive their vaccination. This has been implemented in South Africa for the COVID-19 vaccine: the government has sent out SMS's reminding people to get vaccinated, and a self registration portal has been set up for citizens to enroll in the Electronic Vaccination Data System (EVDS^[This portal can be seen as a type of commitment device, in addition to being a data collection mechanism.]) [@evds]. 

Behavioural economic theory and empirical studies suggest that lotteries can be a useful device for public health interventions. A lottery system can be a cost-effective mechanism for changing behaviour compared to direct transfers because people tend to overweight small probabilities, which is a key insight from the seminal work by @prospect [p.286] on prospect theory. Due to this nonlinear probability weighting, an individual overestimates her chances of winning a lottery. @hiv ran an experiment in Lesotho, where participants were entered into a financial lottery and they could win a cash prize if they tested negative for sexually transmitted infections. In spite of the fact that the lottery had low expected payments, HIV incidence decreased by 21.4\% over two years as a result of the intervention. The study found that the lottery was particularly effective at targeting participants who were more prone to risky sexual behaviour [@hiv]. This supports the theory that risk-seeking individuals value lotteries more. 

The concepts of loss aversion, reference dependence and regret avoidance can also be included in health interventions through a "regret lottery". @prospect describe loss aversion as a cognitive bias whereby people experience losses as more painful than the pleasure they receive from an equivalent gain. Thus, people are more willing to take on risk in order to avoid a loss, and are less risk-seeking when pursuing gain [@prospect p.268]. Reference dependence follows on from loss aversion and suggests that people define gains and losses relative to a reference point [@ref p.1039]. People are also subject to regret avoidance, where there is a significant emotional cost attached to regret and people make decisions to avoid regretting alternative decisions in the future [@regret]. 

A regret lottery takes advantage of these three principles by entering all participants into a lottery but the winner can only claim the prize contingent on some condition. If this condition is not met, a new winner is selected. By entering all participants, people's reference point is shifted to "I have a chance at winning the lottery". However, if a person is not eligible to claim the prize because he does not meet the required condition, he feels he is losing the chance to claim the prize. He is more likely to try and meet the condition in order to minimise the pain of this loss. Additionally, he will want to avoid the regret that would come from having missed the opportunity to claim the prize. 

Several empirical papers investigate how "regret lotteries" can improve health behaviours. @adhere analysed the effect of a daily regret lottery on cholesterol-lowering, and heart medication adherence. Every day, members of the treatment group chose a two-digit number. If a participant's number matched at least one digit of the randomly drawn lottery number, she was eligible to win a cash prize. However, she could only claim her winnings if she had taken her relevant prescribed medication the previous day. The results show that the treatment group better adhered to their medication regime than the control group. In a different study, @regr found that implementing a weekly electronic regret lottery increased adherence to a self-monitoring study protocol. 

Some states in America have attempted to run regret lotteries to encourage people to get vaccinated against COVID-19. An unpublished paper by Thaler _et al_ evaluates the effect of weekly regret lotteries in Philadelphia [@duck]. After giving away \$400,000 in cash prizes to residents, the authors did not find convincing evidence that the regret lotteries significantly increased first-dose vaccination rates for the treatment groups^[The authors acknowledge that there were some design flaws in the experiment that could be clouding the results.]. Similarly disappointing results were found  for a COVID-19 vaccine in Ohio [@ohio]. To the best of my knowledge, there have been no studies on COVID-19 vaccine lotteries in South Africa^[It should be noted, however, that First National Bank is running a COVID-19 vaccine lottery for FNB customers, with total cash prizes amounting to R18 million] - a gap which this experiment is intended to fill. This experiment would also contribute to the empirical literature on lotteries as public health interventions and the integration of behavioural economics into public policy.

# The South African Context \label{context}

An integrated report by @lotto on the lottery habits shows that around 35% of South Africans^[Based on a sample of 3,090 households randomly distributed across the country.] participated in lottery activities in the year preceding the study [@lotto p.78-79]. This suggests that a significant amount of South Africans engage in lotteries, which supports the implementation of a vaccine lottery, as there is evidence of "demand" for lotteries. The average amount spent on lottery activities was R156 per month, whereas the average monthly winnings of lottery schemes was R110. A rational economic agent would recognize that the expected value for playing lotteries is negative ($-$R46) and not buy lottery tickets. The fact that people still play the lotto despite a negative expected value suggests that South Africans may not be rational, and behavioural economic insights are applicable^[The gross revenue from gambling activities, excluding the National Lottery, amounted to R32.7 billion for the 2019 financial year [@gamble p.3]. This represents 0.64\% of South Africa's 2019 nominal GDP (R5.1 trillion) [@statsa p.8], which suggests that gambling is a lucrative market in South Africa and a lotto device could be an appropriate tool for incentivising behaviour.].

Johnson & Johnson (J&J), Pfizer and AstraZeneca are all approved COVID-19 vaccines in South Africa [@sah]. COVID-19 vaccines are available free of charge, but citizens have to register on the EVID portal before receiving their dose^[Due to supply constraints, vaccines were rolled out in phases and are currently available to individuals over the age of 18 only.]. The J&J vaccine only requires 1 dose; the Pfizer and AstraZeneca vaccines require two doses each. For this experiment, _fully vaccinated_ refers to an individual who has had the maximum required doses of any vaccine. _Vaccinated_ refers to an individual who has had 1 shot of any vaccine. According to the @stat, 34% of South Africans are vaccinated, while 25% are fully vaccinated. These low numbers indicate that an intervention is necessary to improve vaccination rates if South Africa wants to achieve herd immunity. Table \ref{tab1} below gives the breakdown of vaccination rates. 

```{r tab1, results = 'asis', warning=FALSE, message=FALSE, echo=FALSE}

library(xtable)
source("code/vac_stats1.R")
data<- vac_stats1()

table <- xtable(data, caption = "Vaccination Statistics \\label{tab1}")
  print.xtable(table,
             # tabular.environment = "longtable",
             floating = TRUE,
             table.placement = 'H',
             # scalebox = 0.3,
             comment = FALSE,
             caption.placement = 'bottom',
             include.rownames = FALSE,
             booktabs = T # Aesthetics
             )

```

# Experiment Design \label{design}

## Sample

The sample used will be the same sample used for the Coronavirus Rapid Mobile Survey (NIDS-CRAM). NIDS-CRAM was created in response to the pandemic as a way to build a representative data set of the South African population to inform decision-making [@nids]. Thus far, there have been five waves of NIDS-CRAM surveys, with wave 5 comprising a sample of 5,862 people being surveyed during April to May 2021 [@nids p.14]. Wave 3 comprised 8,157 potential participants, of which 6,130 were interviewed. For this field experiment, the 8,157 people from Wave 3 will be contacted and asked to participate. We can expect between 5,500 -6,200 people to participate, given the previous rates of attrition in NIDS-CRAM. 

If we assume a conservative sample size of 5,500 and a vaccination proportion of 34\%, then there will be 3630 eligible participants for the study, since we are interested in unvaccinated individuals. This sample will be split into 4 groups of equal size (907), and randomised such that each group has a similar distribution of participants in terms of age, race, health status and gender, as proposed by @random. This randomisation serves to ensure groups are comparable, and eliminate bias in treatment assignments. One group is randomly chosen to be the control group, where individuals will not be entered into any vaccine lottery. The other 3 groups are randomly allocated to different lottery treatments, which are explained below (\ref{group}). There will be 1 lottery a month for each treatment group, which will run (in parallel) for three months. There will be 9 lotteries in total, with 3 lotteries every month. Lottery winners receive a cash prize of R1,000,000. 

## Treatment Groups \label{group}

For the first treatment group, if an individual has received a vaccination shot within a given month, she will be entered into that month's vaccine lottery. At the end of the month, a winner is randomly selected from the lottery pool. Once it is verified that the winner has been vaccinated, she will be privately contacted and will receive a cash prize. The second lottery is a regret lottery. Every individual in the group is entered into a monthly lottery but an individual may only claim her prize if she has been vaccinated in that month. At month end, a winner is randomly drawn from the lottery pool. If the winner has been vaccinated, she will be privately notified and receives a cash prize. If the winner has not been vaccinated, she will receive a "regret message" stating that she would have won the cash prize if she had been vaccinated. 

The final treatment is a "referral lottery". An individual is entered into the monthly lottery if 2 conditions are met: he is vaccinated, and he refers a friend to get vaccinated and the friend gets vaccinated. Both the individual from the sample and his friend are entered into the lottery. At the end of the month, a winner is selected, and once he and his referral partner are verified to be vaccinated, he will be privately informed. An individual can refer more than friend to be vaccinated in any given month. However, only individuals from the treatment group can refer friends (i.e. people outside of group 3 will not be entered into the lottery for referring others to get vaccinated). Group 3 participants may not refer any person in the control group or in group 1 or group 2 (to avoid contamination).

Whenever a lottery is won, it will be announced via SMS to all participants in the relevant treatment group. The amount of the lottery prize and the winner's province will also be included in the SMS. This serves as a reminder of how large the cash prize is, and including the winner's province makes winning seem more tangible. Group 2's SMS will include a reminder that only vaccinated individuals are eligible to win the lottery. Group 3's SMS will include a reminder that participants can refer as many friends as they like to be eligible for the following month's lottery. All SMS's will end with: "Thank you for vaccinating and keeping our country safe!" as one final nudge to encourage/guilt participants to vaccinate.

Table \ref{tab2} summarises the different treatments arms.

```{r tab2, results = 'asis', warning=FALSE, message=FALSE, echo=FALSE}

library(xtable)
source("code/experiments.R")
data<- experiments()

table <- xtable(data, caption = "Treatment Summary \\label{tab2}")
  print.xtable(table,
             # tabular.environment = "longtable",
             floating = TRUE,
             table.placement = 'H',
             # scalebox = 0.3,
             comment = FALSE,
             caption.placement = 'bottom',
             include.rownames = FALSE,
             booktabs = T # Aesthetics
             )

```

## Theory of Change

The decisions and actions associated with getting vaccinated appear deceptively simple but are the result of a complex series of behaviours [@decide p.154]. To design an intervention to increase vaccinations, we need to understand who would be affected by the intervention. There are two groups for whom any treatment is irrelevant: people who would always choose to get vaccinated (regardless of extra incentives), and those who would never get vaccinated. We are interested in identifying a group who would not get vaccinated without intervention but with treatment would get vaccinated. It is plausible that a standard lottery intervention would be appealing to risk-on individuals [@hiv] and could convince them to vaccinate. Even though the vaccine is free in South Africa, poorer individuals may still be reluctant to incur the frictional costs of getting vaccinated (such as transport costs). A lottery intervention could compensate for such costs and incentivise vaccination, especially if such individuals overweight small probabilities (as explained in section \ref{lit}). 

If there are individuals who are willing to get vaccinated but procrastinate (e.g. naïve hyperbolic discounters), a lottery deadline could increase the utility of getting the vaccine earlier enough to overcome the procrastination problem. Additionally, if there are South Africans who are not opposed to the vaccine but expect to free-ride on herd immunity, an incentive in the form of a lottery could way them to get vaccinated. As discussed in the literature review, a regret lottery (treatment 2) would incentivise individuals who are experience reference dependence, loss aversion and regret avoidance to get vaccinated. We would expect vaccination rates to be higher under treatment 2 compared to treatment 1 since a regret lottery has all the same qualities of a standard lottery, and additional incentives. 

Treatment 3 uses aligned incentives^[I like to think of treatment 3 as a positive-sum pyramid scheme] and social pressure to encourage vaccinations. Both the friend and the referring individual get a payoff from being vaccinated because of the lottery entry; however, neither can enter the lottery without the other getting vaccinated, which creates the pressure for each to get vaccinated as soon as possible to avoid disappointing the other. Vaccination incentives are now also a function of the social preferences of group 3 and their friends. It is reasonable to assume that the introduction of a lottery would not deter someone from getting vaccinated, who would have got vaccinated otherwise. Thus, we would expect to see higher vaccination rates for all three treatment groups compared to the control group.

# Treatment and Data \label{treatment}

The two main partner institutions for this field experiment would be the Department of Health, and the CRAM team^[The CRAM project is supported by the Department of Planning Monitoring and Evaluation (DPME), the Research on Socioeconomic Policy (RESEP) group at Stellenbosch University, and the Southern Africa Labour and Development Research Unit (SALDRU) at UCT.]. Vaccine data is collected by health professionals at pharmacies, clinics and other vaccination sites around South Africa. The EVDS captures a vaccinee's personal information^[Names, Identity Number, medical aid details, residential address, email address, phone numbers, employment details, professional category, and health status.] as well as the vaccine date and details. The only additional information needed for the experiment is the name, ID number, and phone number of the referring individual (from treatment group 3). The person receiving the vaccine is responsible for supplying these details to the vaccine administrator. An SMS will be sent to the friend and the referring individual, asking each to confirm that the referal details are correct. The agent who collected the information of the winning person from group 3's lottery will receive R10,000. This is an incentive for healthcare workers to correctly capture referral details.     

The other data collection mechanism occurs through the NIDS-CRAM survey [@quest]. Two extra questions will be added to the original survey: "How often do you buy lotto tickets?" and "How often do you participate in gambling activities?". An additional survey section is needed to record the personal information of the referred friends for group 3. This acts as a second check for the referral information collected by the healthcare worker. It is assumed that the marginal cost of adding this section and the two questions to the survey is negligible. This experiment is designed to harness existing data collecting processes to reduce costs and exploit the existing NIDS-CRAM time series data. 

Funding for the experiment would need to cover: R9,000,000 for the cash prizes, the admin fees associated with the SMS program and award disbursement, and R30,000 for incentive fees for the people who capture the winner's information from group 3's lottery. Potential funding sources include the COVID-19 Africa Rapid Grant Fund administered by the NRF, government funding, and donors such as the Allan & Gill Gray Philanthropy, FEM Education Foundation, and the Michael & Susan Dell Foundation who sponsor the CRAM project.

# Pre-analysis plan \label{pre}

First a balance test 


This is possible because random assignment assures observable baseline characteristics of treatment groups should be similar, statistical unsignificant. This is balance of covariates:

E[X(1)]=E[X(0)]

Where Xn,k

Additionally, if k is sufficiently big. One can end-up in a multiple hypothesis problem, where approximately x% (level of significance) is significant by chance.

Hence, we can test balance by running the following regression:


Then, we can check the balance in covariates by running the F-test of these model against the NULL model $Pr(eachtreat) = +$. If this test is not significant, we can assure that all the covariates X′ are no good to predict the treatment status. Hence, covariates are balanced between treatment groups.

There are several analysis approaches that can be used to analyse the data from this experiment

Start by doing balance test to show randomisation was successful
Compare means of final outcome across treatment groups
This is the most NB result
Run OLS regression to test significance of effect, while controlling for baseline covariates
Also 2SLS if imperfect compliance
Maybe probit / logit if outcome is binary

Estimate effect of treatment on intermediate outcomes
Check whether consistent with theory of change
Estimate interaction between baseline attributes
Check whether subgroup heterogeneity is consistent  with theory of change
Cluster standard errors if multiple observations per cluster or per unit of observation
external validity


Regression discontinuity design

•	a pre-analysis plan of the empirical analysis that will be performed on the data.

https://jamanetwork.com/journals/jama/fullarticle/2781792 - method for analysis



\hfill

<!-- hfill can be used to create a space, like here between text and table. -->


# Conclusion \label{con}
If the experiment results provide enough evidence that a vaccine lottery is effective, the EVDS  nudge is to rollout to whole country and possible  


\newpage

# References {-}

<div id="refs"></div>


