---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

# title: "Helping You Write Academic Papers in R using Texevier"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: TRUE
Entry1: "COVID-19 Vaccine Lottery Field Experiment"
Entry2: "\\textbf{Cassandra Pengelly | 20346212}" # textbf for bold
Entry3: "Professor R. Burger"
Uni_Logo: Tex/logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
Logo_width: 0.9 # If using a logo - use this to set width (size) of image
Entry4: "Behavioural Economics 871 Essay"
Entry5: "15 October 2021"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
# AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
# Author1: "Nico Katzke^[__Contributions:__  \\newline _The authors would like to thank no institution for money donated to this project. Thank you sincerely._]"  # First Author - note the thanks message displayed as an italic footnote of first page.
# Ref1: "Prescient Securities, Cape Town, South Africa" # First Author's Affiliation
# Email1: "nfkatzke\\@gmail.com" # First Author's Email address
# 
# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE
# 
# # Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: TRUE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
# abstract: |
  # Abstract to be written here. The abstract should not be too long and should provide the reader with a good understanding what you are writing about. Academic papers are not like novels where you keep the reader in suspense. To be effective in getting others to read your paper, be as open and concise about your findings here as possible. Ideally, upon reading your abstract, the reader should feel he / she must read your paper in entirety.
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
Example_data <- Texevier::Ex_Dat

# Notice that as you are working in a .Rproj file (I am assuming you are) - the relative paths of your directories start at your specified root.
# This means that when working in a .Rproj file, you never need to use getwd() - it is assumed as your base root automatically.
write_rds(Example_data, path = "data/Example_data.rds")

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->
\newpage

# Introduction \label{Introduction}

Coronavirus disease 2019 (COVID-19) has caused the largest public health crisis, and economic disaster of the 21st century so far [@bad p.471]. According to the @who, COVID-19 has directly resulted in 4,859,277 deaths worldwide^[As of 13 October 2021], and its impact on the global economy has been severe [@bank]. Policymakers are in urgent need of evidence-based strategies to contain the pandemic. As @immun[p.583] reports, one critical mechanism through which epidemics are controlled is through herd immunity. Herd immunity arises when a sufficiently large proportion of the population achieves individual immunity to an infectious disease such that the transmission chain of the disease is halted [@bad p.471]. One method of establishing herd immunity is through vaccination programs [@immun p.583]. Preliminary empirical evidence shows that infection detection and vaccination strategies could be critical tools in subduing COVID-19, for example the study by @erad finds COVID-19 vaccines to be effective in Jakarta, Indonesia. Leaders of other countries have also implemented vaccine programs and the @who reports that a total of 6,364,021,792 vaccine doses have been administered globally as of 10 October 2021.

In line with the international approach, South Africa has opted to issue vaccines in addition to social distancing and lockdown measures. The South African government has stated that it aims to have 67% of the population vaccinated by the end of 2021 [@herd]. However, many South Africans are hesitant to get the COVID vaccine; @stat records that only 25% of South Africa's adult population has been fully vaccinated as of 11 October 2021. @cram compiled a report, based on data from wave 5 of the NIDS-CRAM survey, and find that that around 20% of South Africans are concerned that COVID-19 vaccines are not safe. The report also shows that relatively few people in the sample registered to be vaccinated within two months after registration opened. @cram conclude that a significant portion of South Africans still have to be convinced to take the vaccine^[A sentiment report from @dep shows an interesting break down of different communities beliefs surrounding the vaccine]. While there are different ways to encourage vaccine 

Here, insights from behaviourial economics could provide a cost-effective solution: a vaccine lottery. 






This essay proposes a field experiment to investigate whether a vaccine lottery could improve vaccination rates in South Africa. The experiment explores the effect of three different lottery types - standard, regret and referral - on the take up of vaccines. This essay is structured as follows. Section \ref{lit} briefly reviews the relevant literature on behavioural economics and health incentives. Section \ref{context} elaborates on the South African context. Section \ref{design} describes the design of the experiment and outlines the three types of treatment groups. Section \label{treat} discusses how the treatment will be administered and how the data will be collected.  Lastly, section \ref{pre} gives a pre-analysis plan of the empirical analysis that will be performed on the data, and the final section (\ref{con}) concludes. 

•	a clear statement of the research question and motivation for why this is interesting and important;
•	a brief review of the relevant literature (both theoretical and empirical) which highlights the research gap your experiment will address;

•	a clear description of the experimental design and the theory of change;
•	an explanation of how the treatments will be administered and data gathered (including proposed partner institutions);
•	a pre-analysis plan of the empirical analysis that will be performed on the data.

amount of people have received the vaccine so far. Critical to reach herd immunity is improving vaccination rates. 

# Behavioural Economics and Health Incentives  \label{lit} 

It is not unprecedented to use lotteries in public health interventions. To the best of my knowledge, this experiment would be the first randomized field experiment to assess the impact of lotteries on COVID-19 vaccine take-up in South Africa. Several papers show that regret lotteries can improve health outcomes, such as @regr and @adhere

Loss aversion, regret avoidance, social preferences referral incentives

Individuals have a tendency to overweight small probabilities and this overestimate their chances of winning a lottery.
behavioural economics theory and 

health incentive applications


This paper addresses the gap pf vaccine lotteries in SA.

# The South African Context \label{context}

There are currently four vaccines which have been approved by the South African Health Products Regulatory Authority for use in South Africa: Johnson & Johnson (J&J), Pfizer, Sinovac and AstraZeneca. The two main vaccines being rolled out in South Africa are the J&J and Pfizer, which are both available for free. Vaccines are currently only available to individuals over the age of 18. The J&J vaccine only requires 1 dose; and the Pfizer vaccine requires two doses, 6 weeks apart. For the purposes of this field experiment, _fully vaccinated_ refers to an individual who has had either 1 J&J shot, or both shots of the Pfizer vaccine. _Vaccinated_ refers to an individual who has had 1 shot of either vaccine.

According to @stat, 34% of South Africans are vaccinated, while only 25% are fully vaccinated. 

```{r tab1, results = 'asis', warning=FALSE, message=FALSE, echo=FALSE}

library(xtable)
source("code/vac_stats1.R")
data<- vac_stats1()

table <- xtable(data, caption = "Vaccination Statistics \\label{tab1}")
  print.xtable(table,
             # tabular.environment = "longtable",
             floating = TRUE,
             table.placement = 'H',
             # scalebox = 0.3,
             comment = FALSE,
             caption.placement = 'bottom',
             include.rownames = FALSE,
             booktabs = T # Aesthetics
             )

```

gambling in SA 

# Experiment Design \label{design}

The field experiment is designed to address the research question: could a vaccine lottery improve vaccination rates in South Africa? 

The sample used will be the same sample used for the Coronavirus Rapid Mobile Survey (NIDS-CRAM). NIDS-CRAM was created in response to the pandemic as a way to build a representative data set of the South African population to inform decision-making [@nids]. Thus far, there have been five waves of NIDS-CRAM surveys, with wave 5 comprising a sample of 5,862 people being surveyed during 6 April to 11 May 2021 (@nids p.14). Wave 3 consisted of a total of 8,157 potential participants, of which 6,130 were successfully interviewed. For the purposes of this field experiment, the 8,157 people from Wave 3 will be contacted and asked to participate in the experiment. It seems reasonable to expect between 5,500 and 6,200 people to participate, given the previous rates of attrition experienced in NIDS-CRAM. 

The NIDS-CRAM sample will then be split into 4 groups of equal size. Assuming a conservative sample size of 5,500, each of the 4 groups will comprise 1,375 individuals. @random which will be randomised using the randomisation technique as proposed by @. The first group is the control group, where individuals will not be entered into any vaccine lottery. The other 3 groups are will receive different lottery treatments, which are explained below. There will be 3 monthly lotteries for each treatment group, which will run simultaneously. Thus, in total there will be 9 lotteries for this field experiment, with 3 lotteries every month. Each lottery will award a cash prize of R1,000,000. 

NIDS-CRAM is a special follow-up survey of a subsample of adults from households that were part
of the last wave (2017) of the National Income Dynamics Study (NIDS). NIDS was a large-scale
longitudinal survey, run by the Southern Africa Labour and Development Research Unit (SALDRU),
that tracked the social and economic well-being of South Africans from 2008 up to 2017.
SALDRU (based at the University of Cape Town) was responsible for the NIDS-CRAM survey data
collection, quality assurance and production. The NIDS-CRAM survey instrument includes a wide
range of questions on income and employment, sociodemographic characteristics, and household
welfare. This paper draws on questions about vaccine acceptance that were included in Waves 4
and 5 of the NIDS-CRAM survey. Wave 4 was conducted from 2 February to 10 March 2021 with a
sample of 4,792 individuals, and Wave 5 was conducted from 6 April to 11 May 2021 with a sample
of 4,996. Compared to NIDS-CRAM Wave 1 (May and June 2020), Wave 4 had 31% attrition and
Wave 5 had 28% attrition.

For the first treatment group, if an individual has received a vaccination shot^[An individual only need a receive one shot of any vaccine - receiving the first shot of the Pfizer qualifies an individual for that month's lottery] within a given month, she will be entered into that month's vaccine lottery. At the end of the month, a winner is randomly selected from the lottery pool. Once it is verifed that the winner has been vaccinated, she will be privately contacted and will receive a cash prize. It will then be announced via sms to everyone in the treatment group (group 1) that the lottery has been won, the amount of the lottery prize, and the winner's province. 

Following the approach of @regret, the second lottery is a "regret lottery". Every individual in the sample is entered into a monthly lottery but an individual may only claim her prize if she has been vaccinated (i.e. had at least 1 shot of any vaccine). At month end, a winner is randomly drawn from the lottery pool. If the winner has been vaccinated, she will be privately notified and will receive a cash prize. It will then be announced via sms to everyone in the treatment group (group 3) that the lottery has been won, the amount of the cash prize, the winner's province and a reminder that only vaccinated individuals are eligible to win the lottery.

The final treatment is a "referral lottery". An individual is entered into the monthly lottery if 2 conditions are met: he is vaccinated, and he refers a friend to get vaccinated and the friend gets vaccinated. Both the individual from the sample and his friend are entered into the lottery. At the end of the month, a winner is selected, verified to be vaccinated and privately informed. An SMS is sent out to the treatment group with the same information as for group 2 and group 3, in addition to a message thanking the sample for protecting their friends as well as a reminder that they can refer more friends to be eligible for the following month's lottery. 

Including an SMS after the lottery has won serves as a reminder that there is a 

Table \ref{tab2} summarises the different treatments administered.


```{r tab2, results = 'asis', warning=FALSE, message=FALSE, echo=FALSE}

library(xtable)
source("code/experiments.R")
data<- experiments()

table <- xtable(data, caption = "Treatment Summary \\label{tab2}")
  print.xtable(table,
             # tabular.environment = "longtable",
             floating = TRUE,
             table.placement = 'H',
             # scalebox = 0.3,
             comment = FALSE,
             caption.placement = 'bottom',
             include.rownames = FALSE,
             booktabs = T # Aesthetics
             )

```

Many South Africans are hesitant to get the COVID vaccine. As STATSA shows, amount of people have received the vaccine so far. Critical to reach herd immunity is improving vaccination rates. 
In order to improve the take-up of vaccinations, a field experiment designed around a vaccination lottery is proposed. While some governments have considered and experimented with lump-sum payments, behavioural economics could provide a more cost-effective solution. Individuals have a tendency to overweight small probabilities and this overestimate their chances of winning a lottery. In South Africa, 

Theory: overweight small probabilities, gambling, social preferences, regret avoidance
Empirical: vaccine field designs, lottery incentives, regret lottery incentives
Several authors have experimented with lotteries as an incentive for vaccinations, although there have been no studies as of yet on the South African population.
Probles with vaccine studies: too few participants. A larger study could fill this gap in the literature.


# Treatment and Data \label{treat}

The CRAM survey exists to provide monthly nationally-representative data on key outcomes such as unemployment, household income, child hunger and access to government grants.



Partner with NIDS, department of health, vaccine administer, funding
data collected at vaccine 
add 3 questions to NIDS data cram

# Pre-analysis plan \label{pre}

Students must submit an essay in which they design a field experiment that could answer an interesting behavioural economic question. The essay must contain the following:

•	a clear statement of the research question and motivation for why this is interesting and important;
•	a brief review of the relevant literature (both theoretical and empirical) which highlights the research gap your experiment will address;

•	a clear description of the experimental design and the theory of change;
•	an explanation of how the treatments will be administered and data gathered (including proposed partner institutions);
•	a pre-analysis plan of the empirical analysis that will be performed on the data.

Overview
In this field experiment, a person who refers his/her friend to receive a vaccine would be entered into a lucky draw, with a monetary prize, created by the government. The purpose behind this nudge is to encourage people who would otherwise not have got a Covid vaccine, to do so. The hypothesis is that there should be an increase in the total number of people receiving a Covid vaccine after the nudge is implemented. Increasing the number of vaccinations is important as medical research shows that vaccines decrease the probability of contracting Covid-19 and are also effective at reducing the severity of the symptoms of the virus for those who do contract it. 
The Nudge
The nudge addresses behaviour by creating an environment where there is social pressure to get a vaccine (if I wanted to enter the lucky draw, I would pressure my friend into getting the vaccine). It is also likely that if a person asks her friend to get the vaccine so she can enter the lucky draw, she will reciprocate and get the vaccine as well so that her friend may enter the draw, which will also increase the number of people getting vaccinated. 
For the vaccines that require two doses (e.g. Pfizer), a person’s name could be withdrawn, if the second shot is not given within a certain amount of time. This makes use of loss aversion, where people who already have their names in the draw feel the pain of having their names withdrawn more intensely than the pleasure of having their names added a second time to the draw for getting their second shot.
Target Group
The lucky draw is anticipated to attract people who are risk-on (they enjoy gambling, and are less worried about getting vaccinated), and poorer individuals for whom winning money is more attractive. These target groups are desirable as they are less likely to get the vaccine, and the government would like to maximise the number of vaccinated people. Additionally, if there are individuals who want to be vaccinated but procrastinate getting the vaccine (e.g. naïve hyperbolic discounters), setting a deadline for the lucky draw could increase the utility of getting the vaccine earlier enough to overcome the procrastination problem. There is no downside or extra cost for having people enter the lucky draw who would otherwise still have got the vaccine.
Proposed Partner Institutions
This field experiment would be in collaboration with the South African government and facilities that conduct vaccinations (e.g. Clicks). The government would be where the data is centralized and the administers of vaccines would all be data collection nodes. After a person has received a vaccine, the administer would ask if the person received a referral for the shot, and then note the ID number of the friend in addition to the individual’s details. 

Data Collection
There is a data collection system already set up at the vaccination sites so this extra data point would not be difficult to collect within the current tracking system. Depending on costs, the referral friend could be sent an sms thanking her for caring about others and getting them vaccinated, and letting her know that she has been entered into the draw. This is a positive reinforcement technique and shows people that the government is following up on their promise. This acknowledgement and transparency is expected to encourage more referrals. Once the lucky draw has been concluded, the data can be analysed, the purpose of which is to uncover whether the nudge increased vaccinations. 

> I suggest renaming the top line after \@article, as done in the template ref.bib file, to something more intuiti

```{r Figure1,  warning =  FALSE, fig.align = 'center', fig.cap = "Caption Here \\label{Figure1}", fig.ext = 'png', fig.height = 3, fig.width = 6}

  # This is just a random plot to show you a plot. This is done if the getwd() does not point to your Template's directory.
  # Set WD below to your project's working directory.

  g <-
  ggplot(data = mtcars) +
  geom_point(aes(x = disp, y = hp, color = cyl)) +
  theme_bw() +
  # theme(legend.position = "none") +
  theme(plot.title = element_text(size = 10)) +
  theme(axis.title = element_text(size = 10),
        axis.title.x = element_text()) +
  labs(y = "y-axis", x = "x-axis", title = "Some amazing plot")

g

```



To reference calculations __in text__, _do this:_ From table \ref{tab1} we see the average value of mpg is `r mean(mtcars[1:5,]$mpg)`.

```{r LongTable, results = 'asis'}

library(xtable)

data = mtcars %>% tibble::as_tibble()
  addtorow          <- list()
  addtorow$pos      <- list()
  addtorow$pos[[1]] <- c(0)
  addtorow$command  <- c(paste("\\hline \n",
                               "\\endhead \n",
                               "\\hline \n",
                               "{\\footnotesize Continued on next page} \n",
                               "\\endfoot \n",
                               "\\endlastfoot \n",sep=""))
table <- xtable(data, caption = "Long Table Example")
  print.xtable(table,
             tabular.environment = "longtable",
             floating = FALSE, # Leave this as is.
             table.placement = 'H', # Leave this as is.
             booktabs = T, # Aesthetics
             include.rownames = FALSE,  # Typically you don't want this in a table.
             add.to.row = addtorow, # For adding the Continued on next page part...
             comment = FALSE,
             caption.placement = 'top',  # Where do you want the caption?
             size="\\fontsize{12pt}{13pt}\\selectfont"  # Size of text in table..
             )
# See https://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf for table inspiration
```

\hfill

<!-- hfill can be used to create a space, like here between text and table. -->


# Conclusion \label{con}

I hope you find this template useful. Remember, stackoverflow is your friend - use it to find answers to questions. Feel free to write me a mail if you have any questions regarding the use of this package. To cite this package, simply type citation("Texevier") in Rstudio to get the citation for @Texevier (Note that uncited references in your bibtex file will not be included in References).

<!-- Make title of bibliography here: -->
<!-- \newpage -->

\newpage

# References {-}

<div id="refs"></div>


# Appendix {-}

## Appendix A {-}

Some appendix information here

## Appendix B {-}

